# Introduction to Data Science HW 4
##### Copyright Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva


```{r}
# Enter your name here: Shaun Mckellar Jr
```

### Attribution statement: (choose only one and delete the rest)


```{r}
# 2. I did this homework with help from the book and the professor and these Internet sources:Youtube
```

Reminders of things to practice from previous weeks: <br>
Descriptive statistics: 	mean( ) max( ) min( )<br>
Coerce to numeric:		as.numeric( )<br>

## Part 1: Use the Starter Code 

Below, I have provided a starter file to help you. <br>

Each of these lines of code **must be commented** (the comment must that explains what is going on, so that I know you understand the code and results).


```{r}
library(tidyverse)
library(dplyr)
#Decoding JSON
library(jsonlite)
#URL OF JSON Data
dataset <- url("https://intro-datascience.s3.us-east-2.amazonaws.com/role.json")
#Converts the JSON data into a structured R object
readlines <- jsonlite::fromJSON(dataset)
#Used to access nested elements within the readlines object
df <- readlines$objects$person
```

A.	Explore the **df** dataframe (e.g., using head() or whatever you think is best).


```{r}
head(df)
df
```

B.	Explain the dataset <br>
o	What is the dataset about? <br>
o	How many rows are there and what does a row represent?<br>
o	How many columns and what does each column represent?<br>



```{r}
#To find the details about the dataset using the summary function and you get the details of the dataset.
summary(df)
#The row reps the records in the dataset
nrow(df)
#To find the number of columns in a dataset using the ncold function
ncol(df)
#To find a columns in a dataset using the colnames function
colnames(df)
```

C. What does running this line of code do? Explain in a comment:


```{r}
#The substr function is for get substring in data or a vector. If you look at the birthday it contains the values like "1951-05-20"
vals <- substr(df$birthday,1,4)
```

D. Create a new attribute 'age' - how old the person is 
**Hint:** You may need to convert it to numeric first.


```{r}
#The vals variable contains the year and we have to convert it using the as.numeric function
old <- as.numeric(vals)
#Create a new variable age with value as the current year and then subtracts with the birth year from the old variable to get the age.
age <- 2021-old
age
```

E. Create a function that reads in the role json dataset, and adds the age attribute to the dataframe, and returns that dataframe


```{r}
#We are having to add the age to the dataset df using the data.frame function.
df <- data.frame(df,age)
head(df)
```

F. Use (call, invoke) the function, and store the results in df


```{r}
#Importing the dplyr library using the library function
library(dplyr)
df%>%filter(gender=='female')%>%nrow()
```

## Part 2: Investigate the resulting dataframe 'df'

A.	How many senators are women? 


```{r}
#Importing the dplyr library using the library function
library(dplyr)
women <- sum(df$gender == "female")
women
```

B.	How many senators have a YouTube account?  


```{r}
#Importing the dplyr library using the library function
library(dplyr)
df%>%filter(youtubeid != 'NA')%>%nrow()
```

C.	 How many women senators have a YouTube account?


```{r}
library(dplyr)
df%>%filter(gender=='female' & youtubeid !='NA')%>%nrow()
```

D.	Create a new dataframe called **youtubeWomen** that only includes women senators who have a YouTube account.   


```{r}
library(dplyr)
youtubeWomen <- df%>%filter(gender=='female' & youtubeid != 'NA')
youtubeWomen
```

E.	Make a histogram of the **age** of senators in **youtubeWomen**, and then another for the senetors in **df**. Add a comment describing the shape of the distributions.


```{r}
#Importing the dplyr library using the library function
library(dplyr)
#To make the a histogram we have to use the hist function. We pass the age column from the df data uing the $ operator, and xlab for label x-axis.
hist(youtubeWomen$age,xlab = 'Age')
```
```{r}
#Importing the dplyr library using the library function
library(dplyr)
#To make the a histogram we have to use the hist function. We pass the age column from the df data uing the $ operator, and xlab for label x-axis.
hist(df$age,xlab = 'Age')
```

It seems the age of the senators are between 60-70. Most of the senators are older than 60 as well. The most common age for Women that are senators are in their 60s.