---
output:
  html_document: default
  pdf_document: default
---
# Intro to Data Science - HW 6
##### Copyright Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva


```{r}
# Enter your name here: Shaun McKellar Jr
```


### Attribution statement: (choose only one and delete the rest)


```{r}
# 1. I did this homework by myself, with help from the book and the professor.
# 3. I did this homework with help from Angela Ozelski but did not cut and paste any code.
```

```{r}
library(dplyr)
library(dbplyr)
library(tidyverse)
library("ggplot2")
library(tinytex)
library(jsonlite)
library("rjson")
library("maps")
library("ggmap")
library("mapproj")
```

Last assignment we explored **data visualization** in R using the **ggplot2** package. This homework continues to use ggplot, but this time, with maps.  In addition, we will merge datasets using the built-in **merge( )** function, which provides a similar capability to a **JOIN in SQL** (don't worry if you do not know SQL). Many analytical strategies require joining data from different sources based on a **“key”** – a field that two datasets have in common. 

## Step 1: Load the population data

A.	Read the following JSON file, https://intro-datascience.s3.us-east-2.amazonaws.com/cities.json and store it in a variable called **pop**.

Examine the resulting pop dataframe and add comments explaining what each column contains.



```{r}
pop <- jsonlite::fromJSON("https://intro-datascience.s3.us-east-2.amazonaws.com/cities.json")
```

```{r}
#Lists the growth from 2000-13
#The rank based on population
#The state and city
#The lat and long
```

B.	Calculate the **average population** in the dataframe. Why is using mean() directly not working? Find a way to correct the data type of this variable so you can calculate the average (and then calculate the average)

Hint: use **str(pop)** or **glimpse(pop)** to help understand the dataframe


```{r}
#Needed to change the type to numeric
mean(as.numeric(pop$population))
str(pop)
```

C.	What is the population of the smallest city in the dataframe? Which state is it in?


```{r}
#finds the index
which.min(pop$population)
```
```{r}
#Florida, Panama City 36877 people
pop[1000,]
```

## Step 2: Merge the population data with the state name data

D)	Read in the state name .csv file from the URL below into a dataframe named **abbr** (for "abbreviation") – make sure to use the read_csv() function from the tidyverse package: <br>
https://intro-datascience.s3.us-east-2.amazonaws.com/statesInfo.csv



```{r}
read_csv(url('https://intro-datascience.s3.us-east-2.amazonaws.com/statesInfo.csv')) -> abbr
```

E)	To successfully merge the dataframe **pop** with the **abbr** dataframe, we need to identify a **column they have in common** which will serve as the **“key”** to merge on. One column both dataframes have is the **state column**. The only problem is the slight column name discrepancy – in **pop**, the column is called **“state”** and in **abbr** – **“State.”** These names need to be reconciled for the merge() function to work. Find a way to rename **abbr’s “State”** to **match** the **state column in pop**.    


```{r}
abbr <- abbr %>%
  rename(state = State)
```

F)	Merge the two dataframes (using the **‘state’ column** from both dataframes), storing the resulting dataframe in **dfNew**.


```{r}
dfNew <- merge(abbr,pop, by = "state")
```

G)	Review the structure of **dfNew** and explain the columns (aka attributes) in that dataframe.


```{r}
#Lists the growth from 2000-2013
#The rank based on population
#The state, city and abbreviation
#The lat and long
```

## Step 3: Visualize the data

H)	Plot points (on top of a map of the US) for **each city**. Have the **color** represent the **population**.


```{r}
states <- map_data("state")
ggplot(NULL) +
  geom_polygon(data = states, color="black", fill="white",
               aes(x=long,y=lat,group=group)) + coord_map() +
  geom_point(data=dfNew,aes(x=longitude,y=latitude,color=as.numeric(population)))
```
1: Yes

I)	Add a block comment that criticizes the resulting map. It’s not very good.


```{r}
#Seems very clustered and hard to tell what's going on in the graph
```

## Step 4: Group by State


J)	Use group_by and summarise to make a dataframe of state-by-state population. Store the result in **dfSimple**.


```{r}
dfSimple <- dfNew %>% 
  group_by(state) %>% 
  summarize(population_state = sum(as.numeric(population)))
```

K)	Name the most and least populous states in **dfSimple** and show the code you used to determine them.


```{r}
which.min(dfSimple$population_state)
```
```{r}
#Vermont is the last populus state
dfSimple[46,1]
```
```{r}
which.max(dfSimple$population_state)
```
```{r}
dfSimple[5,1]
```
```{r}
#California is the most populus state
```

## Step 5: Create a map of the U.S.,  with the color of the state representing the state population

L) Make sure to expand the limits correctly and that you have used **coord_map** appropriately.



```{r}
dfSimple$region <- tolower(dfSimple$state) 

theUS <- merge(states,dfSimple, region = region) 

theUS %>% ggplot() + geom_polygon(aes( x= long, y= lat, group= group,, fill = theUS$population)) + coord_map()
```
