---
title: "Intro to Data Science - HW 10 - Shiny"
output: html_document
runtime: shiny
---

##### Copyright Jeffrey Stanton, Jeffrey Saltz, Christopher Dunham, and Jasmina Tacheva

```{r}
# Enter your name here: Shaun McKellar Jr
```


### Attribution statement: (choose only one and delete the rest)

```{r}
# 1. I did this homework by myself, with help from the book and the professor.
```

At a high level, we want to create a scatter plot, where the user can select what is the x axis and the y axis of the scatter plot. We will use the same dataset as used in our first ggplot homework.

The size and color will be represented by the year attribute, and the user will be able to choose one of the following (for  the x and the y axis):
new_sp_m014
new_sp_f014
new_sp_m65
new_sp_f65


```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(shinyFiles)
library(shiny)
library(curl)
```

## Read in the same dataset we used for the ggplot homework

The file is: "https://intro-datascience.s3.us-east-2.amazonaws.com/who.csv" and store it in the *tb* dataframe
```{r}
tb <- read_csv("https://intro-datascience.s3.us-east-2.amazonaws.com/who.csv", show_col_types = FALSE) 
```
## Clean up the dataset, just like we did in the ggplot homework.

First, remove na's from iso2
```{r}
tb = subset(tb, is.na(tb$iso2) == FALSE )
```


Now create the dataframe 'tbCan', which is the iso2 for canada (CA)
```{r}
tbCan <- subset(tb,tb$iso2 == 'CA')
```

We will need the imputeTS package (only install if needed)
```{r}
#install.packages('imputeTS')
library(imputeTS)
```

Now we can use 'na_interpolation for new_sp_m014
```{r}
tbCan$new_sp_m014 <- na_interpolation(tbCan$new_sp_m014)
```


## Create the User Interface of our shiny app

Define the sidebarPanel, which is two choices (use 'selectInput'), one for the x axis of the scatter plot, and the other is the y axis for the scatter plot. (make sure to library shiny).
```{r}

inputChoices <- c(
      "males aged 0-14 years" = "new_sp_m014",
      "females aged 0-14 years" = "new_sp_f014",
      "males aged 65 years and over" = "new_sp_m65",
      "females aged 65 and over" = "new_sp_f65")
sidePanel <- sidebarPanel(
    #your code goes here
     selectInput("variable1", "X-axis:",inputChoices),
  selectInput("variable2", "Y-axis:", inputChoices),width = 10   )

```

Create the mainPanel to show the scatter plot.

```{r}
mainPanel <-  mainPanel(
    # your code goes here
       plotOutput("distplot"))
```

# Define UI for application (called ui)

Use a 'fluidPage' using the sidebarLayout, and your sidePanel and mainPanel
```{r}
ui1 <- fluidPage(

    # Application title
    titlePanel("Tuberculosis cases in Canada"),

    # Sidebar with a slider input for number of bins 
    sidebarLayout(

         #add the slider
      sidebarPanel(
        sliderInput("Bins", "Number of bins:", min = 0, max = 1000, value = 500),sidePanel),
      plotOutput("plot"))
)
```

## Now let's define the server

Use ggplot to render a scatter plot, using the tbCan dataframe, the input for the x-axis and the input for the y-axis.

Store the results in the 'server' variable

```{r}
server2 <- function(input, output) {

  tbCan <- readtb()
  output$plot <- renderPlot(
    ggplot(tbCan, aes(x=tbCan[,input$variable1], y=tbCan[,input$variable2])) 
    + ylab(input$variable2) + xlab(input$variable1)
    + geom_point(aes(size = year, color= year )) + stat_summary_bin(fun.y = 'mean',bins = input$Bins,color = 'orange',size = 2, geom = 'point')
  ) 
}
  
  #is the same as the one previously defined
readtb <- function() {
  #cleanup
  testFrame <- read.csv(curl("https://intro-datascience.s3.us-east-2.amazonaws.com/who.csv")) 
  testFrame = subset(testFrame, is.na(testFrame$iso2) == FALSE )
  testFrame <- subset(testFrame,testFrame$iso2 == 'CA')
  testFrame$new_sp_m014 <- na_interpolation(testFrame$new_sp_m014)
  
  #convert columns to numbers
 
  testFrame$new_sp_m014 <-Numberize(testFrame$new_sp_m014)
  testFrame$new_sp_f014 <-Numberize(testFrame$new_sp_f014)
  testFrame$new_sp_m65 <-Numberize(testFrame$new_sp_m65)
  testFrame$new_sp_f65 <-Numberize(testFrame$new_sp_f65)
  
  rownames(testFrame) <- NULL
  return(testFrame)
}

Numberize <- function(inputVector)
{
  # Get rid of commas
  inputVector<-gsub(",","", inputVector)
  # Get rid of spaces
  inputVector<-gsub(" ","", inputVector)
  return(as.numeric(inputVector))
}

```

Now run the shiny App
```{r}
# Run the application 
shinyApp(ui = ui1, server = server2)
```
