---
output:
  html_document: default
  pdf_document: default
---
# Intro to Data Science - HW 5
##### Copyright Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva

```{r}
# Enter your name here: Shaun McKellar Jr
```


### Attribution statement: (choose only one and delete the rest)


```{r}
# 1. I did this homework by myself, with help from the book and the professor.
# 2. I did this homework with help from the book and the professor and these Internet sources:https://community.rstudio.com/
```

**This module:** **Data visualization** is important because many people can make sense of data more easily when it is presented in graphic form. As a data scientist, you will have to present complex data to decision makers in a form that makes the data interpretable for them. From your experience with Excel and other tools, you know that there are a variety of **common data visualizations** (e.g., pie charts). How many of them can you name? <br>

The most powerful tool for data visualization in R is called **ggplot**. Written by computer/data scientist **Hadley Wickham**, this **“graphics grammar”** tool builds visualizations in layers. This method provides immense flexibility, but takes a bit of practice to master.


## Step 1: Make a copy of the data

A.	Read the **who** dataset from this URL: https://intro-datascience.s3.us-east-2.amazonaws.com/who.csv <br>
into a new dataframe called **tb**. 

Your new dataframe, tb, contains a so-called **multivariate time series**: a sequence of measurements on 23 Tuberculosis-related (TB) variables captured repeatedly over time (1980-2013). Familiarize yourself with the nature of the 23 variables by consulting the dataset’s codebook which can be found here: https://intro-datascience.s3.us-east-2.amazonaws.com/TB_data_dictionary_2021-02-06.csv.

```{r}
library(dbplyr)
library(tidyverse)
library("ggplot2")
library(readxl)
library(tinytex)
```

```{r}
library(imputeTS)
```

```{r}
tb <- read.csv('https://intro-datascience.s3.us-east-2.amazonaws.com/who.csv')
```

B.	How often were these measurements taken (in other words, at what frequency were the variables measured)? Put your answer in a comment. 


```{r}
observations <- tb %>% group_by(year) %>% tally() 
observations
summary(observations)
```

## Step 2: Clean-up the NAs and create a subset

A.	Let's clean up the iso2 attribute in **tb** 

Hint: use *is.na()* -- well use *! is.na()* 


```{r}
tb <- tb[!(is.na(tb$iso2)), ]
```

B. Create a subset of **tb** containing **only the records for Canada ("CA" in the iso2 variable)**. Save it in a new dataframe called **tbCan**. Make sure this new df has **29 observations and 23 variables**.


```{r}
tbCan <- tb[tb$iso2 == 'CA',]
```

C. A simple method for dealing with small amounts of **missing data** in a numeric variable is to **substitute the mean of the variable in place of each missing datum**. <br>
This expression locates (and reports to the console) all the missing data elements in the variable measuring the **number of positive pulmonary smear tests for male children 0-4 years old** (there are 26 data points missing)


```{r}
tbCan$new_sp_m04[is.na(tbCan$new_sp_m04)]
```


    Error in eval(expr, envir, enclos): object 'tbCan' not found
    Traceback:



D.	Write a comment describing how that statement works.


```{r}
#this expression is returns each row in new_sp_04 that has an NA present in the row
```

E.	Write 4 more statements to check if there is missing data for the number of positive pulmonary smear tests for: **male and female** children 0-14 years old (**new_sp_m014** and **new_sp_f014**), and **male and female citizens 65 years of age and older**, respectively. What does empty output suggest about the number of missing observations?


```{r}
youngmales <- tbCan$new_sp_m014[is.na(tbCan$new_sp_m014)]
youngmales
```

```{r}
youngfemales <- tbCan$new_sp_f014[is.na(tbCan$new_sp_f014)]
youngfemales
```


```{r}
oldermales <- tbCan$new_sp_m65[is.na(tbCan$new_sp_m65)]
oldermales
```

```{r}
olderfemales <- tbCan$new_sp_f65[is.na(tbCan$new_sp_f65)]
olderfemales
```

```{r}
# We received integer(0) for all of these dataframes we created because there is no missing data for these attributes
```

There is an R package called **imputeTS** specifically designed to repair missing values in time series data. We will use this instead of mean substitution.  <br>
The **na_interpolation()** function in this package takes advantage of a unique characteristic of time series data: **neighboring points in time can be used to “guess” about a missing value in between**.

F.	Install the **imputeTS** package (if needed) and use **na_interpolation( )** on the variable from part C. Don’t forget that you need to save the results back to the **tbCan** dataframe. Also update any attribute discussed in part E (if needed).


```{r}
library('imputeTS')
```
```{r}
tbCan$new_sp_m04 <- na_interpolation(tbCan$new_sp_m04)
tbCan$new_sp_m014 <- na_interpolation(tbCan$new_sp_m014)
tbCan$new_sp_m65 <- na_interpolation(tbCan$new_sp_m65)
tbCan$new_sp_f014 <- na_interpolation(tbCan$new_sp_f014)
tbCan$new_sp_f65 <- na_interpolation(tbCan$new_sp_f65)
```

G.	Rerun the code from C and E above to check that all missing data have been fixed. 


```{r}
males04 <- tbCan$new_sp_m04[is.na(tbCan$new_sp_m04)]
males04
```


```{r}
youngmales <- tbCan$new_sp_m014[is.na(tbCan$new_sp_m014)]
youngmales
```

```{r}
youngfemales <- tbCan$new_sp_f014[is.na(tbCan$new_sp_f014)]
youngfemales
```

```{r}
oldermales <- tbCan$new_sp_m65[is.na(tbCan$new_sp_m65)]
oldermales
```

```{r}
olderfemales <- tbCan$new_sp_f65[is.na(tbCan$new_sp_f65)]
olderfemales
```

## Step 3: Use ggplot to explore the distribution of each variable
**Don’t forget to install and library the ggplot2 package.** Then: <br>
H.	Create a histogram for **new_sp_m014**. Be sure to add a title and briefly describe what the histogram means in a comment.


```{r}
library(ggplot2)
hist(tbCan$new_sp_m014, main = "Histogram for Positive cases Males 14 Yrs old and younger", xlab = "Cases")
```

I.	Create histograms (using ggplot) of each of the other three variables from E with ggplot( ). <br>
Which parameter do you need to adjust to make the other histograms look right?


```{r}
tbCan %>% ggplot() +  
  geom_histogram(binwidth = 1, fill="blue", color="white", 
                 aes(x=new_sp_m014)) +
                  ggtitle('Histogram for Positive Cases Males 14y or younger in CA')

```

```{r}
tbCan %>% ggplot() +  
  geom_histogram(binwidth = 1, fill="blue", color="white", 
                 aes(x=new_sp_f014)) +
                  ggtitle('Histogram for Positive Cases Females 14y or younger in CA')
```
```{r}
tbCan %>% ggplot() +  
  geom_histogram(binwidth = 1, fill="blue", color="white", 
                 aes(x=new_sp_m65)) +
                  ggtitle('Histogram for Positive Cases Male 65+ in CA')
```

```{r}
tbCan %>% ggplot() +  
  geom_histogram(binwidth = 1, fill="blue", color="white", 
                 aes(x=new_sp_f65)) +
                  ggtitle('Histogram for Positive Cases Female 65+ in CA')
```

## Step 4: Explore how the data changes over time
J.	These data were collected in a period of several decades (1980-2013). You can thus observe changes over time with the help of a line chart. Create a **line chart**, with **year** on the X-axis and **new_sp_m014** on the Y-axis.


```{r}
tbCan %>% ggplot() +  
  geom_line(color = 'black',
                 aes(x=year, y=new_sp_m014)) +
                  ggtitle('Line Graph for Positive Cases Males 14y or younger in CA')
```

K.	Next, create similar graphs for each of the other three variables. Change the **color** of the line plots (any color you want).


```{r}
tbCan %>% ggplot() +  
  geom_line(color = 'green', 
                 aes(x=year, y=new_sp_f014)) +
                  ggtitle('Line Graph for Positive Cases Females 14y or younger in CA')
```
```{r}
tbCan %>% ggplot() +  
  geom_line(color = 'red', 
                 aes(x=year, y=new_sp_m65)) +
                  ggtitle('Line Graph for Positive Cases Males 65+ in CA')
```
```{r}
tbCan %>% ggplot() +  
  geom_line(color = 'orange', 
                 aes(x=year, y=new_sp_f65)) +
                  ggtitle('Line Graph for Positive Cases Females 65+ in CA')
```

L.	Using vector math, create a new variable by combining the numbers from **new_sp_m014** and **new_sp_f014**. Save the resulting vector as a new variable in the **tbCan** df called **new_sp_combined014**. This new variable represents the number of positive pulmonary smear tests for male AND female children between the ages of 0 and 14 years of age. Do the same for SP **tests among citizens 65 years of age and older** and save the resulting vector in the tbCan variable called **new_sp_combined65**.


```{r}
tbCan$new_sp_combined014 <- tbCan$new_sp_m014 + tbCan$new_sp_f014
tbCan$new_sp_combined65 <- tbCan$new_sp_m65 + tbCan$new_sp_f65
```

M.	Finally, create a **scatter plot**, showing **new_sp_combined014** on the x axis, **new_sp_combined65** on the y axis, and having the **color and size** of the point represent **year**.


```{r}
tbCan %>% ggplot() +  geom_point() +  aes(x=new_sp_combined014, y=new_sp_combined65, size=year, color=year ) +  ggtitle('Scatter Plot of Positive Test for Male & Female with the ages 0-14 and 65 +')

```

N.	Interpret this visualization – what insight does it provide?


```{r}
#The older people have more tb test than the younger population through the years, with  the number of positive cases beginning to decline
```
