---
output:
  pdf_document: default
  html_document: default
---

## Intro to Data Science - HW 1

##### Copyright Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva

```{r}
# Enter your name here: Shaun Mckellar Jr
```

### Attribution statement: (choose only one and delete the rest)

```{r}
# 1. I did this homework by myself, with help from the book and the professor.
```

Define a variable:

```{r}
x <- 280
```

Define the following vectors, which represent the **population** (in thousands) and **number of colleges** in each of the five counties in Central New York (CNY) -- **Cayuga**, **Cortland**, **Madison**, **Onondaga**, and **Oswego**, in this order:

```{r}
population <- c(80, 49, 73, 467, 122)
colleges <- c(2, 2, 3, 9, 2)
```

# Part 1: Calculating statistics using R

A.  Show the number of observations in the **population** vector with the length() function:

```{r}
length(population)
```

B.  Show the number of observations in the **colleges** vector with the length() function:

```{r}
length(colleges)
```

C.  Calculate the average CNY population using the mean() function:

```{r}
mean(population)
```

D.  Calculate the average number of colleges in CNY using the mean() function:

```{r}
mean(colleges)
```

E.  Calculate the total CNY population using the sum() function:

```{r}
sum(population)
```

F.  Calculate the total number of colleges in CNY using the sum() function:

```{r}
sum(colleges)
```

G.  Calculate the average CNY population again, this time using **the results from steps A & E**:

```{r}
sum(population)/length(population)
```

H.  Calculate the average number of colleges in CNY again, this time using **the results from steps B & F**:

```{r}
sum(colleges)/length(colleges)
```

# Part 2: Using the max/min and range functions in {r}

I.  How many colleges does the county with most colleges have? Hint: Use the max() function:

```{r}
max(colleges)
```

J.  What is the population of the least populous county in CNY? **Hint:** Use the min() function:

```{r}
min(population)
```

K.  Display the populations of the least populous and most populous county in the dataset together. **Hint:** Use the range() function:

```{r}
range(population)
```

# Part 3: Vector Math

L.  Create a new vector called **extraPop**, which is the current population of a county **+ 50** (each county has 50,000 more people):

```{r}
extraPop <- population +50
```

M.  Calculate the average of **extraPop**:

```{r}
mean(extraPop)
```

N.  In a variable called **bigCounties**, store all the population numbers from the original **population** vector which are **greater than 120** (using **subsetting** in R):

```{r}
bigCounties <- population[population > 120]
```

O.  Report the length of **bigCounties**:

```{r}
length(bigCounties)
```
